\documentclass[12pt,a4paper]{book}

\usepackage{morewrites}

\usepackage{import}
\usepackage[sort]{standalone}

\usepackage{graphicx,color,psfrag}
\usepackage{tabularx,colortbl}

\usepackage{epsfig}


%\usepackage{listings}
%\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage[colorlinks=true]{hyperref}   % use for hypertext links, including those to external documents and URLs

\usepackage{scalefnt}

%\usepackage{theorem}

\usepackage{tabu}
%\usepackage{forest}
\usepackage{cleveref}
\usepackage{xspace}
\usepackage[backend=bibtex,sorting=ydnt,firstinits=true,maxbibnames=99]{biblatex}

\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}

\usepackage{pgfplots}
\pgfplotsset{compat=newest}

\usepackage{placeins}


% number chapters in each part separately
\makeatletter
\@addtoreset{chapter}{part}
\makeatother


\addbibresource{bibliography.bib}
\addbibresource{wpg.bib}
\addbibresource{pepper.bib}

\input{texsetup.tex}

\newcommand{\x}{\V{\chi}}
\newcommand{\objb}{b}
\newcommand{\violation}{\V{v}}
\newcommand{\objA}{\M{\mathcal{A}}}


\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}
\let\cleardoublepage\clearpage


\title{\projectname}
\author{}
\date{\today}


\begin{document}
\begin{refsection}
\maketitle
\tableofcontents

\clearpage
\phantomsection
%\pdfbookmark[1]{List of Acronyms}{listofacronyms}
\printacronyms
\addcontentsline{toc}{chapter}{List of Acronyms}


\part{Core}

\chapter{Introduction}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is Humoto}\label{sec.intro}

\projectname{} -- is a software framework for manipulation of linear
    least-squares problems with (in)equality constraints. It supports both
    weighting and lexicographic prioritization and can be characterized as a
    tool for goal programming \cite{wiki2017gp}. However, the development was
    driven by works in other fields -- robotics, control, and numerical
    optimization, \EG, \cite{Dimitrov2015preprint, Saab2012tranrob,
    Escande2014ijrr}; for this reason our terminology and interpretations are
    different.


The core functionalities of \projectname are formulation of least-squares
    problems and their resolution using various third-party software. Both of
    these operations are performed through unified \ac{API}. Moreover, due to
    our interest in robotic applications, the framework facilitates formulation
    and implementation of optimization problems for control of robots, in
    particular, \ac{MPC} problems. For the same reason, we pay special
    attention to computational performance in order to be able to employ the
    framework in real-time scenarios.


In addition to the core components, the distribution of \projectname includes
    several modules -- implementations of specific controllers. The modules
    serve as examples of using the framework, but can also be used in
    accordance with their primary purpose. You can learn more about the
    provided modules in \cref{part.wpg} and \cref{part.pepper}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Outline}\label{sec.intro}

The present document introduces basic concepts behind the \projectname and
    contains various mathematical derivations which are used in the framework
    and the modules. However, this document cannot serve as a comprehensive
    guide -- you have to refer to \sn{Doxygen} documentation in order to learn
    how to use \projectname in your applications or learn more about
    implementation details.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Notation}\label{sec.notation}

\begin{description}
    \item[Software names] \hfill \\
        Names of programs and software libraries, names of constants, variables
        and functions that are used in programs are typed in a monospaced font:
        \sn{Eigen}.

    \item[Scalars, vectors, matrices] \hfill
        \begin{itemize}
            \item Vectors and matrices are denoted by letters in a bold font:
                $\V{v}$, $\M{M}$, $\M{\mathcal{A}}$.

            \item Scalars are denoted using the standard italic or calligraphic
                font: $N, n, \mathcal{K}$.

            \item $\T{(\cdot)}$ -- transpose of a matrix or a vector.

            \item $\CROSS[(\cdot)]$ -- a skew-symmetric matrix used for
                representation of a cross product of two three dimensional
                vectors as a product of a matrix and a vector:
                %
                \begin{equation}
                    \V{v}
                    =
                    \begin{bmatrix}
                        v^x\\
                        v^y\\
                        v^z
                    \end{bmatrix},
                    \quad
                    \CROSS[\V{v}]
                    =
                    \begin{bmatrix}
                        0      &   -v^z   &   v^y\\
                        v^z    &   0      &   -v^x\\
                        -v^y   &   v^x    &   0 \\
                    \end{bmatrix}.
                \end{equation}

            \item Block diagonal matrices:
                %
                \begin{equation}
                    \begin{gathered}
                        \diag{2}{\M{M}} =
                        \begin{bmatrix}
                            \M{M}   &   \M{0}\\
                            \M{0}   &   \M{M}\\
                        \end{bmatrix}
                        ,
                        \quad
                        \diag{k = 1 \dots 2}{\M{M}_k} =
                        \begin{bmatrix}
                            \M{M}_1 &   \M{0}  \\
                            \M{0}   &   \M{M}_2\\
                        \end{bmatrix}
                        ,
                        \\
                        \diag{}{\M{M}, \M{R}} =
                        \begin{bmatrix}
                            \M{M} &   \M{0}  \\
                            \M{0}   &   \M{R}\\
                        \end{bmatrix}
                        .
                    \end{gathered}
               \end{equation}

            \item Stacked vectors and matrices:
                %
                \begin{equation}
                    \V{v} = (\V{v}_1, \dots, \V{v}_n) = \begin{bmatrix} \V{v}_1 \\ \vdots \\ \V{v}_n \\ \end{bmatrix},
                    \quad
                    \M{M} = (\M{M}_1, \dots, \M{M}_n) = \begin{bmatrix} \M{M}_1 \\ \vdots \\ \M{M}_n \\ \end{bmatrix}.
                \end{equation}

            \item Inequalities between vectors $\V{v} \ge \V{r}$ are interpreted
                component-wise.
        \end{itemize}


    \item[Special matrices and vectors] \hfill
        \begin{itemize}
            \item $\M{I}$ -- an identity matrix. $\M{I}_n$ -- $n \CROSS n$ identity matrix.

            \item $\M{I}_{(\cdot)}$ -- a selection matrix.

            \item $\M{0}$ -- a matrix of zeros. $\M{0}_{n,m}$ -- $n \CROSS m$ matrix of zeros.
        \end{itemize}


    \item[Reference frames] \hfill
        \begin{itemize}
            \item Frames are denoted using a sans-serif font: $\FRAME{A}$. All
                considered frames are orthonormal.

            \item $\V[A]{v}$ -- vector expressed in frame $\FRAME{A}$.

            \item $\M[B][A]{R}$ -- rotation matrix from from frame $\FRAME{B}$
                to frame $\FRAME{A}$. For example, 2-d rotation matrix is defined
                as
                %
                \begin{equation}
                    \M[B][A]{R}
                    =
                    \begin{bmatrix}
                        \cos(\theta) &   -\sin(\theta)\\
                        \sin(\theta) &   \cos(\theta)\\
                    \end{bmatrix},
                \end{equation}
                %
                where $\theta$ is the rotation angle of frame $\FRAME{B}$ with
                respect to $\FRAME{A}$.

            \item The global frame is implicit and is not denoted by any
                letter, \EG, $\M[B][]{R}$ rotates from frame $\FRAME{B}$ to the
                global frame.
        \end{itemize}


    \item[Sets] \hfill
        \begin{itemize}
            \item The sets are denoted using a blackboard bold font: $\SET{A}$.

            \item $\RR$ is the set of real numbers.

            \item $\RR_{\ge 0}, \RR_{> 0}$ are the sets of non-negative and positive real numbers.

            \item $\RR^n$ is the set of real-valued vectors.

            \item $\RR^{n \CROSS m}$ is the set of real-valued matrices.
        \end{itemize}


    \item[Other] \hfill
        \begin{itemize}
            \item Function names in mathematical expressions are written in the
                regular font: $\FUNC{func}(\V{x}, \V{y})$.

            \item $\NORME{\cdot}$ denotes the Euclidean norm.
        \end{itemize}
\end{description}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{optimization}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{reference}


\printbibliography[heading=bibintoc, title=Bibliography]
\end{refsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% modules
\input{wpg}
\input{pepper.tex}


%\appendix
\end{document}
